---
title: "AM6"
author: "SA"
date: "12/03/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries


```{r}
library(tidyverse)
library(reshape2)
library(ggthemes)
library(ggrepel)
library(RColorBrewer)
library(ChannelAttribution)
library(markovchain)
library(visNetwork)
library(expm)
library(stringr)
library(purrrlyr)
```

# Generate and prepare data


```{r}
set.seed(99669966)
df_raw <- data.frame(customer_id = paste0('id', sample(c(1:20000), replace = TRUE)), date = as.Date(rbeta(80000, 0.7, 10) * 100, origin = "2016-01-01"), channel = paste0('channel_', sample(c(0:7), 80000, replace = TRUE, prob = c(0.2, 0.12, 0.03, 0.07, 0.15, 0.25, 0.1, 0.08))) ) %>%
  group_by(customer_id) %>%
  mutate(conversion = sample(c(0, 1), n(), prob = c(0.975, 0.025), replace = TRUE)) %>%
  ungroup() %>%
  dmap_at(c(1, 3), as.character) %>%
  arrange(customer_id, date)

df_raw


df_raw <- df_raw %>%
  mutate(channel = ifelse(channel == 'channel_2', NA, channel))



df_paths <- df_raw %>%
  group_by(customer_id) %>%
  mutate(path_no = ifelse(is.na(lag(cumsum(conversion))), 0, lag(cumsum(conversion))) + 1) %>%
  ungroup()

df_paths
```

## For first purchaser only


```{r}


df_paths_1 <- df_paths %>%
  filter(path_no == 1) %>%
  select(-path_no)

df_paths_1
```

# replace some channels

```{r}


##### replace some channels #####
df_path_1_clean <- df_paths_1 %>%
  # removing NAs
  filter(!is.na(channel)) %>%
  
  # adding order of channels in the path
  group_by(customer_id) %>%
  mutate(ord = c(1:n()),
         is_non_direct = ifelse(channel == 'channel_6', 0, 1),
         is_non_direct_cum = cumsum(is_non_direct)) %>%
  
  # removing Direct (channel_6) when it is the first in the path
  filter(is_non_direct_cum != 0) %>%
  
  # replacing Direct (channel_6) with the previous touch point
  mutate(channel = ifelse(channel == 'channel_6', channel[which(channel != 'channel_6')][is_non_direct_cum], channel)) %>%
  
  ungroup() %>%
  select(-ord, -is_non_direct, -is_non_direct_cum)



df_path_1_clean
```






```{r}
df_path_1_clean <- df_path_1_clean %>%
  group_by(customer_id) %>%
  mutate(uniq_channel_tag = ifelse(length(unique(channel)) == 1, TRUE, FALSE)) %>%
  ungroup()


df_path_1_clean_multi <- df_path_1_clean %>%
  filter(uniq_channel_tag == FALSE) %>%
  select(-uniq_channel_tag)

```


# computing time lapses from the first contact to conversion/last contact

```{r}
# computing time lapses from the first contact to conversion/last contact
df_multi_paths_tl <- df_path_1_clean_multi %>%
  group_by(customer_id) %>%
  summarise(path = paste(channel, collapse = ' > '),
            first_touch_date = min(date),
            last_touch_date = max(date),
            tot_time_lapse = round(as.numeric(last_touch_date - first_touch_date)),
            conversion = sum(conversion)) %>%
  ungroup()
```

# distribution plot
```{r}

ggplot(df_multi_paths_tl %>% filter(conversion == 1), aes(x = tot_time_lapse)) +
  theme_minimal() +
  geom_histogram(fill = '#4e79a7', binwidth = 1)
```


# cumulative distribution plot

```{r}
ggplot(df_multi_paths_tl %>% filter(conversion == 1), aes(x = tot_time_lapse)) +
  theme_minimal() +
  stat_ecdf(geom = 'step', color = '#4e79a7', size = 2, alpha = 0.7) +
  geom_hline(yintercept = 0.95, color = '#e15759', size = 1.5) +
  geom_vline(xintercept = 23, color = '#e15759', size = 1.5, linetype = 2)
```

# data for
1) time lapse from the first contact,

2) time lapse between the conversion date and a previous contact.

```{r}
#1) time lapse from the first contact,

#2) time lapse between the conversion date and a previous contact.


df_multi_paths_tl_2 <- df_path_1_clean_multi %>%
  group_by(customer_id) %>%
  mutate(prev_touch_date = lag(date)) %>%
  ungroup() %>%
  filter(conversion == 1) %>%
  mutate(prev_time_lapse = round(as.numeric(date - prev_touch_date)))

# distribution
ggplot(df_multi_paths_tl_2, aes(x = prev_time_lapse)) +
  theme_minimal() +
  geom_histogram(fill = '#4e79a7', binwidth = 1)

# cumulative distribution
ggplot(df_multi_paths_tl_2, aes(x = prev_time_lapse)) +
  theme_minimal() +
  stat_ecdf(geom = 'step', color = '#4e79a7', size = 2, alpha = 0.7) +
  geom_hline(yintercept = 0.95, color = '#e15759', size = 1.5) +
  geom_vline(xintercept = 12, color = '#e15759', size = 1.5, linetype = 2)
```

#subsetting data for For tot_time_lapse > 20 & prev_touch > 10

```{r}
#For tot_time_lapse > 20 & prev_touch > 10


df_multi_paths_tl_3 <- df_path_1_clean_multi %>%
  group_by(customer_id) %>%
  mutate(prev_time_lapse = round(as.numeric(date - lag(date)))) %>%
  summarise(path = paste(channel, collapse = ' > '),
            tot_time_lapse = round(as.numeric(max(date) - min(date))),
            prev_touch_tl = prev_time_lapse[which(max(date) == date)],
            conversion = sum(conversion)) %>%
  ungroup() %>%
  mutate(is_fruitless = ifelse(conversion == 0 & tot_time_lapse > 20 & prev_touch_tl > 10, TRUE, FALSE)) %>%
  filter(conversion == 1 | is_fruitless == TRUE)

df_multi_paths_tl_3
```

#models for 
#multi-channel paths only for the reporting period (e.g. 90 days) in the example. Therefore, paths include a minimum of 2 touches with 2 dates and last touch date is equal to conversion date.
1) Criteria #1 - 23 days period between those 2 or more dates (between 1st and last touches) covers 95% of customers with conversions.
2) Criteria #2 - 12 days period between last 2 touches (between last and previous touches) covers 95% of customers with conversions.

```{r}
##### Generic Probabilistic Model #####
df_all_paths_compl <- df_multi_paths_tl_3 %>%
  
mutate(null_conversion = ifelse(conversion == 1, 0, 1))

```

#Finding the order of the model
```{r}


res=choose_order(df_all_paths_compl, var_path="path", var_conv="conversion",
                 var_null="null_conversion")
#plot auc and penalized auc
plot(res$auc$order,res$auc$auc,type="l",xlab="order",ylab="pauc",main="AUC")
lines(res$auc$order,res$auc$pauc,col="red")
legend("right", legend=c("auc","penalized auc"),
       col=c("black","red"),lty=1)
## End(Not run)



```

#model with order = 1 and best order
```{r}

mod_attrib_complete <- markov_model(
  df_all_paths_compl,
  var_path = 'path',
  var_conv = 'conversion',
  var_null = 'null_conversion', order = 1,
  out_more = TRUE
)

mod_attrib_complete_best <- markov_model(
  df_all_paths_compl,
  var_path = 'path',
  var_conv = 'conversion',
  var_null = 'null_conversion', order = res$suggested_order,
  out_more = TRUE
)
```


#For Model with order = 1

```{r}
trans_matrix_prob <- mod_attrib_complete$transition_matrix %>%
  dmap_at(c(1, 2), as.character)

trans_matrix_prob

mod_attrib_complete$removal_effects
mod_attrib_complete$result
```

#For Model with order = Best

```{r}
mod_attrib_complete_best$removal_effects
mod_attrib_complete_best$result
```

# Calculate ROAS and CPA - for model with order 1


```{r}
calculation = mod_attrib_complete$result

calculation <- data.frame(total_cost = c(15000, 21000, 22000, 10000, 20000, 5000), calculation)

calculation$chanel_weight <- calculation$total_conversions / sum(calculation$total_conversions)

calculation$cost_weight <- calculation$total_cost / sum(calculation$total_cost)

calculation$roas <- calculation$chanel_weight / calculation$cost_weight

calculation$optimal_budget = calculation$total_cost * calculation$roas

calculation$CPA = calculation$total_cost / calculation$total_conversions



calculation$CPA
calculation$optimal_budget
calculation$total_cost
```


# Create an ordered graph showing conversions attributed to each channel

```{r}
# Create an ordered graph showing conversions attributed to each channel
g_channel_performance <- ggplot(calculation, aes(x = channel_name, y = total_conversions, fill = channel_name)) + 
  geom_bar(stat = "identity", width = 0.6) +
  ylim(0, 350) +
  scale_fill_manual(values = c("#CE2D4F",
                               "#A14DA0",
                               "#9D79BC",
                               "#7F96FF",
                               "#A9CEF4",
                               "#2d35ce")) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 9, angle = 30, hjust = 0.6, face = "bold")) +
  theme(panel.grid.major.x = element_blank()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = round(total_conversions, 0)), fontface = "bold", size = 4, vjust = -1) + 
  labs(x = "", y = "Conversions") +
  ggtitle("Channel Performance") +
  guides(fill=FALSE)

g_channel_performance
```


```{r}
df_g2 = calculation[, c("channel_name", "total_cost", "optimal_budget")]
df_g2 = melt(df_g2, id = "channel_name")

g_budget_allocation <- ggplot(df_g2, aes(x = channel_name, y = value, fill = variable)) + 
  geom_bar(stat = "identity", width = 0.6, position = position_dodge(width = 0.7)) +
  scale_fill_manual(labels = c("Current Budget", "Optimal Budget"), values = c("#FFD166", "#04A777")) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 30, hjust = 0.6, face = "bold")) +
  theme(panel.grid.major.x = element_blank()) +
  geom_text(aes(label = round(value, 0)), 
            fontface = "bold", size = 3.5, 
            vjust = -0.5, position = position_dodge(width = 0.75)) +
  labs(x = "", y = "Budget $") +
  ggtitle("Budget Allocation") +
  theme(plot.title = element_text(hjust = 0.5))

g_budget_allocation
```

# transition matrix heatmap

```{r}
############## visualizations ##############
# transition matrix heatmap for "real" data
df_plot_trans <- mod_attrib_complete$transition_matrix

cols <- c("#e7f0fa", "#c9e2f6", "#95cbee", "#0099dc", "#4ab04a", "#ffd73e", "#eec73a",
          "#e29421", "#e29421", "#f05336", "#ce472e")
t <- max(df_plot_trans$transition_probability)

ggplot(df_plot_trans, aes(y = channel_from, x = channel_to, fill = transition_probability)) +
  theme_minimal() +
  geom_tile(colour = "white", width = .9, height = .9) +
  scale_fill_gradientn(colours = cols, limits = c(0, t),
                       breaks = seq(0, t, by = t/4),
                       labels = c("0", round(t/4*1, 2), round(t/4*2, 2), round(t/4*3, 2), round(t/4*4, 2)),
                       guide = guide_colourbar(ticks = T, nbin = 50, barheight = .5, label = T, barwidth = 10)) +
  geom_text(aes(label = round(transition_probability, 2)), fontface = "bold", size = 4) +
  theme(legend.position = 'bottom',
        legend.direction = "horizontal",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face = "bold", vjust = 2, color = 'black', lineheight = 0.8),
        axis.title.x = element_text(size = 24, face = "bold"),
        axis.title.y = element_text(size = 24, face = "bold"),
        axis.text.y = element_text(size = 8, face = "bold", color = 'black'),
        axis.text.x = element_text(size = 8, angle = 90, hjust = 0.5, vjust = 0.5, face = "plain")) +
  ggtitle("Transition matrix heatmap")
```



# Calculate ROAS and CPA - for model with order = Best

```{r}
calculation = mod_attrib_complete_best$result

calculation <- data.frame(total_cost = c(15000, 21000, 22000, 10000, 20000, 5000), calculation)

calculation$chanel_weight <- calculation$total_conversions / sum(calculation$total_conversions)

calculation$cost_weight <- calculation$total_cost / sum(calculation$total_cost)

calculation$roas <- calculation$chanel_weight / calculation$cost_weight

calculation$optimal_budget = calculation$total_cost * calculation$roas

calculation$CPA = calculation$total_cost / calculation$total_conversions



calculation$CPA
calculation$optimal_budget
calculation$total_cost
```



```{r}
```


# Create an ordered graph showing conversions attributed to each channel

```{r}
# Create an ordered graph showing conversions attributed to each channel
g_channel_performance <- ggplot(calculation, aes(x = channel_name, y = total_conversions, fill = channel_name)) + 
  geom_bar(stat = "identity", width = 0.6) +
  ylim(0, 350) +
  scale_fill_manual(values = c("#CE2D4F",
                               "#A14DA0",
                               "#9D79BC",
                               "#7F96FF",
                               "#A9CEF4",
                               "#2d35ce")) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 9, angle = 30, hjust = 0.6, face = "bold")) +
  theme(panel.grid.major.x = element_blank()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = round(total_conversions, 0)), fontface = "bold", size = 4, vjust = -1) + 
  labs(x = "", y = "Conversions") +
  ggtitle("Channel Performance") +
  guides(fill=FALSE)

g_channel_performance
```


```{r}
df_g2 = calculation[, c("channel_name", "total_cost", "optimal_budget")]
df_g2 = melt(df_g2, id = "channel_name")

g_budget_allocation <- ggplot(df_g2, aes(x = channel_name, y = value, fill = variable)) + 
  geom_bar(stat = "identity", width = 0.6, position = position_dodge(width = 0.7)) +
  scale_fill_manual(labels = c("Current Budget", "Optimal Budget"), values = c("#FFD166", "#04A777")) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 30, hjust = 0.6, face = "bold")) +
  theme(panel.grid.major.x = element_blank()) +
  geom_text(aes(label = round(value, 0)), 
            fontface = "bold", size = 3.5, 
            vjust = -0.5, position = position_dodge(width = 0.75)) +
  labs(x = "", y = "Budget $") +
  ggtitle("Budget Allocation") +
  theme(plot.title = element_text(hjust = 0.5))

g_budget_allocation
```

