---
title: "MMM"
author: "SA"
date: "04/03/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

**Main steps that we will review are the following:**

-- splitting paths depending on purchases counts
-- replacing some channels/touch points
-- a unique channel/touchpoint case
-- consequent duplicated channels in the path and higher order Markov chains
- paths that haven’t led to a conversion
- customer journey duration
- attributing revenue and costs comparisons

```{r cars, echo=FALSE}
library(tidyverse)
library(reshape2)
library(ggthemes)
library(ggrepel)
library(RColorBrewer)
library(ChannelAttribution)
library(markovchain)
library(visNetwork)
library(expm)
library(stringr)
library(purrrlyr)
```




```{r}
##### simulating the "real" data #####
set.seed(454)
df_raw <- data.frame(customer_id = paste0('id', sample(c(1:20000), replace = TRUE)), date = as.Date(rbeta(80000, 0.7, 10) * 100, origin = "2016-01-01"), channel = paste0('channel_', sample(c(0:7), 80000, replace = TRUE, prob = c(0.2, 0.12, 0.03, 0.07, 0.15, 0.25, 0.1, 0.08))) ) %>%
        group_by(customer_id) %>%
        mutate(conversion = sample(c(0, 1), n(), prob = c(0.975, 0.025), replace = TRUE)) %>%
        ungroup() %>%
        dmap_at(c(1, 3), as.character) %>%
        arrange(customer_id, date)
 
df_raw <- df_raw %>%
        mutate(channel = ifelse(channel == 'channel_2', NA, channel))
```




```{r}
head(df_raw)
```

##SPLITTING PATHS DEPENDING ON PURCHASES COUNTS

```{r}
##### splitting paths #####
df_paths <- df_raw %>%
        group_by(customer_id) %>%
        mutate(path_no = ifelse(is.na(lag(cumsum(conversion))), 0, lag(cumsum(conversion))) + 1) %>%
        ungroup()
```

Paths:
1) channel_1 -> conversion #1

2) channel_4 -> channel_0 -> channel_6 -> conversion #2

3) channel_4 -> conversion #3

4) channel_6 -> NA -> channel_5


Using the Life-Cycle Grids concept (or a different one) we can split the data set into different sets and compute attribution separately for different customer segments. For simplicity, we will compute attribution for first-purchasers only.

```{r}
df_paths_1 <- df_paths %>%
        filter(path_no == 1) %>%
        select(-path_no)
```
You can also embed plots, for example:

```{r}
##### replace some channels #####
df_path_1_clean <- df_paths_1 %>%
        # removing NAs
        filter(!is.na(channel)) %>%
         
        # adding order of channels in the path
        group_by(customer_id) %>%
        mutate(ord = c(1:n()),
               is_non_direct = ifelse(channel == 'channel_6', 0, 1),
               is_non_direct_cum = cumsum(is_non_direct)) %>%
         
        # removing Direct (channel_6) when it is the first in the path
        filter(is_non_direct_cum != 0) %>%
         
        # replacing Direct (channel_6) with the previous touch point
        mutate(channel = ifelse(channel == 'channel_6', channel[which(channel != 'channel_6')][is_non_direct_cum], channel)) %>%
         
        ungroup() %>%
        select(-ord, -is_non_direct, -is_non_direct_cum)
```


ONE- AND MULTI-CHANNEL PATHS ISSUE
split a unique channel and multi-channel paths.

Split data for paths with one or more unique channels
Calculate total conversions for one-channel paths and compute the Markov model for multi-channel paths
Summarize results for each channel.

```{r}
##### one- and multi-channel paths #####
df_path_1_clean <- df_path_1_clean %>%
        group_by(customer_id) %>%
        mutate(uniq_channel_tag = ifelse(length(unique(channel)) == 1, TRUE, FALSE)) %>%
        ungroup()
 
df_path_1_clean_uniq <- df_path_1_clean %>%
        filter(uniq_channel_tag == TRUE) %>%
        select(-uniq_channel_tag)
 
df_path_1_clean_multi <- df_path_1_clean %>%
        filter(uniq_channel_tag == FALSE) %>%
        select(-uniq_channel_tag)
 
### experiment ###
# attribution model for all paths
df_all_paths <- df_path_1_clean %>%
        group_by(customer_id) %>%
        summarise(path = paste(channel, collapse = ' > '),
                  conversion = sum(conversion)) %>%
        ungroup() %>%
        filter(conversion == 1)
 
mod_attrib <- markov_model(df_all_paths,
                           var_path = 'path',
                           var_conv = 'conversion',
                           out_more = TRUE)
mod_attrib$removal_effects
mod_attrib$result
d_all <- data.frame(mod_attrib$result)
 
# attribution model for splitted multi and unique channel paths
df_multi_paths <- df_path_1_clean_multi %>%
        group_by(customer_id) %>%
        summarise(path = paste(channel, collapse = ' > '),
                  conversion = sum(conversion)) %>%
        ungroup() %>%
        filter(conversion == 1)
 
mod_attrib_alt <- markov_model(df_multi_paths,
                           var_path = 'path',
                           var_conv = 'conversion',
                           out_more = TRUE)
mod_attrib_alt$removal_effects
mod_attrib_alt$result
 
# adding unique paths
df_uniq_paths <- df_path_1_clean_uniq %>%
        filter(conversion == 1) %>%
        group_by(channel) %>%
        summarise(conversions = sum(conversion)) %>%
        ungroup()
 
d_multi <- data.frame(mod_attrib_alt$result)
 
d_split <- full_join(d_multi, df_uniq_paths, by = c('channel_name' = 'channel')) %>%
        mutate(result = total_conversions + conversions)
 
sum(d_all$total_conversions)
sum(d_split$result)
```


When using one-order Markov chains, a subsequence of the same channels in a path (duplicates) can be reduced to one channel (for example C2 in the path C1 → C2 → C2 → C2 → C3 can be reduced to C1 → C2 → C3). Because, mathematically, it doesn’t matter how many times each C2 goes through the loop with itself in the transition matrix, it will be in the C3 state finally. Therefore, we will obtain different transition matrices but the same Removal Effect for channels with or without subsequent duplicates.

```{r}
##### Higher order of Markov chains and consequent duplicated channels in the path #####
 
# computing transition matrix - 'manual' way
df_multi_paths_m <- df_multi_paths %>%
        mutate(path = paste0('(start) > ', path, ' > (conversion)'))
m <- max(str_count(df_multi_paths_m$path, '>')) + 1 # maximum path length
 
df_multi_paths_cols <- colsplit(string = df_multi_paths_m$path, pattern = ' > ', names = c(1:m))
colnames(df_multi_paths_cols) <- paste0('ord_', c(1:m))
df_multi_paths_cols[df_multi_paths_cols == ''] <- NA
 
df_res <- vector('list', ncol(df_multi_paths_cols) - 1)
 
for (i in c(1:(ncol(df_multi_paths_cols) - 1))) {
         
        df_cache <- df_multi_paths_cols %>%
                select(num_range("ord_", c(i, i+1))) %>%
                na.omit() %>%
                group_by(.dots = c(paste0("ord_", c(i, i+1)))) %>%
                summarise(n = n()) %>%
                ungroup()
         
        colnames(df_cache)[c(1, 2)] <- c('channel_from', 'channel_to')
        df_res[[i]] <- df_cache
}
 
df_res <- do.call('rbind', df_res)
 
df_res_tot <- df_res %>%
        group_by(channel_from, channel_to) %>%
        summarise(n = sum(n)) %>%
        ungroup() %>%
        group_by(channel_from) %>%
        mutate(tot_n = sum(n),
               perc = n / tot_n) %>%
        ungroup()
 
df_dummy <- data.frame(channel_from = c('(start)', '(conversion)', '(null)'),
                       channel_to = c('(start)', '(conversion)', '(null)'),
                       n = c(0, 0, 0),
                       tot_n = c(0, 0, 0),
                       perc = c(0, 1, 1))
 
df_res_tot <- rbind(df_res_tot, df_dummy)
 
# comparing transition matrices
trans_matrix_prob_m <- dcast(df_res_tot, channel_from ~ channel_to, value.var = 'perc', fun.aggregate = sum)
trans_matrix_prob <- data.frame(mod_attrib_alt$transition_matrix)
trans_matrix_prob <- dcast(trans_matrix_prob, channel_from ~ channel_to, value.var = 'transition_probability')
 
# computing attribution - 'manual' way
channels_list <- df_path_1_clean_multi %>%
        filter(conversion == 1) %>%
        distinct(channel)
channels_list <- c(channels_list$channel)
 
df_res_ini <- df_res_tot %>% select(channel_from, channel_to)
df_attrib <- vector('list', length(channels_list))
 
for (i in c(1:length(channels_list))) {
         
        channel <- channels_list[i]
         
        df_res1 <- df_res %>%
                mutate(channel_from = ifelse(channel_from == channel, NA, channel_from),
                       channel_to = ifelse(channel_to == channel, '(null)', channel_to)) %>%
                na.omit()
         
        df_res_tot1 <- df_res1 %>%
                group_by(channel_from, channel_to) %>%
                summarise(n = sum(n)) %>%
                ungroup() %>%
                 
                group_by(channel_from) %>%
                mutate(tot_n = sum(n),
                       perc = n / tot_n) %>%
                ungroup()
         
        df_res_tot1 <- rbind(df_res_tot1, df_dummy) # adding (start), (conversion) and (null) states
         
        df_res_tot1 <- left_join(df_res_ini, df_res_tot1, by = c('channel_from', 'channel_to'))
        df_res_tot1[is.na(df_res_tot1)] <- 0
         
        df_trans1 <- dcast(df_res_tot1, channel_from ~ channel_to, value.var = 'perc', fun.aggregate = sum)
         
        trans_matrix_1 <- df_trans1
        rownames(trans_matrix_1) <- trans_matrix_1$channel_from
        trans_matrix_1 <- as.matrix(trans_matrix_1[, -1])
         
        inist_n1 <- dcast(df_res_tot1, channel_from ~ channel_to, value.var = 'n', fun.aggregate = sum)
        rownames(inist_n1) <- inist_n1$channel_from
        inist_n1 <- as.matrix(inist_n1[, -1])
        inist_n1[is.na(inist_n1)] <- 0
        inist_n1 <- inist_n1['(start)', ]
         
        res_num1 <- inist_n1 %*% (trans_matrix_1 %^% 100000)
         
        df_cache <- data.frame(channel_name = channel,
                               conversions = as.numeric(res_num1[1, 1]))
         
        df_attrib[[i]] <- df_cache
}
 
df_attrib <- do.call('rbind', df_attrib)
 
# computing removal effect and results
tot_conv <- sum(df_multi_paths_m$conversion)
 
df_attrib <- df_attrib %>%
        mutate(tot_conversions = sum(df_multi_paths_m$conversion),
               impact = (tot_conversions - conversions) / tot_conversions,
               tot_impact = sum(impact),
               weighted_impact = impact / tot_impact,
               attrib_model_conversions = round(tot_conversions * weighted_impact)
        ) %>%
        select(channel_name, attrib_model_conversions)
```


compute complete probabilistic model for the first paths we have

```{r}
##### Generic Probabilistic Model #####
df_all_paths_compl <- df_path_1_clean %>%
        group_by(customer_id) %>%
        summarise(path = paste(channel, collapse = ' > '),
                  conversion = sum(conversion)) %>%
        ungroup() %>%
        mutate(null_conversion = ifelse(conversion == 1, 0, 1))
 
mod_attrib_complete <- markov_model(
        df_all_paths_compl,
        var_path = 'path',
        var_conv = 'conversion',
        var_null = 'null_conversion',
        out_more = TRUE
)
 
trans_matrix_prob <- mod_attrib_complete$transition_matrix %>%
        dmap_at(c(1, 2), as.character)
 
##### viz #####
edges <-
        data.frame(
                from = trans_matrix_prob$channel_from,
                to = trans_matrix_prob$channel_to,
                label = round(trans_matrix_prob$transition_probability, 2),
                font.size = trans_matrix_prob$transition_probability * 100,
                width = trans_matrix_prob$transition_probability * 15,
                shadow = TRUE,
                arrows = "to",
                color = list(color = "#95cbee", highlight = "red")
        )
 
nodes <- data_frame(id = c( c(trans_matrix_prob$channel_from), c(trans_matrix_prob$channel_to) )) %>%
        distinct(id) %>%
        arrange(id) %>%
        mutate(
                label = id,
                color = ifelse(
                        label %in% c('(start)', '(conversion)'),
                        '#4ab04a',
                        ifelse(label == '(null)', '#ce472e', '#ffd73e')
                ),
                shadow = TRUE,
                shape = "box"
        )
 
visNetwork(nodes,
           edges,
           height = "2000px",
           width = "100%",
           main = "Generic Probabilistic model's Transition Matrix") %>%
        visIgraphLayout(randomSeed = 123) %>%
        visNodes(size = 5) %>%
        visOptions(highlightNearest = TRUE)
```

Let’s assume that we are going to attract 1000 visits from channel_5 and we want to model how many conversions we will obtain or see what channels customers will make contact with in several steps. The script involves manipulating the matrix of transition probabilities associated with the Markov chain.

Once we have the transition matrix computed, we can project in what state (channel) that customer contact will be, for example, in 5 steps (5th degree) or how many conversion we can obtain (we use 100,000 degrees to makes sure that all customers transited through the transition matrix).


```{r}
##### modeling states and conversions #####
# transition matrix preprocessing
trans_matrix_complete <- mod_attrib_complete$transition_matrix
trans_matrix_complete <- rbind(trans_matrix_complete, df_dummy %>%
                                       mutate(transition_probability = perc) %>%
                                       select(channel_from, channel_to, transition_probability))
trans_matrix_complete$channel_to <- factor(trans_matrix_complete$channel_to, levels = c(levels(trans_matrix_complete$channel_from)))
trans_matrix_complete <- dcast(trans_matrix_complete, channel_from ~ channel_to, value.var = 'transition_probability')
trans_matrix_complete[is.na(trans_matrix_complete)] <- 0
rownames(trans_matrix_complete) <- trans_matrix_complete$channel_from
trans_matrix_complete <- as.matrix(trans_matrix_complete[, -1])
 
 
# creating empty matrix for modeling
model_mtrx <- matrix(data = 0,
                     nrow = nrow(trans_matrix_complete), ncol = 1,
                     dimnames = list(c(rownames(trans_matrix_complete)), '(start)'))
# adding modeling number of visits
model_mtrx['channel_5', ] <- 1000
 
c(model_mtrx) %*% (trans_matrix_complete %^% 5) # after 5 steps
c(model_mtrx) %*% (trans_matrix_complete %^% 100000) # after 100000 steps
```


CUSTOMER JOURNEY DURATION
```{r}
##### Customer journey duration #####
# computing time lapses from the first contact to conversion/last contact
df_multi_paths_tl <- df_path_1_clean_multi %>%
        group_by(customer_id) %>%
        summarise(path = paste(channel, collapse = ' > '),
                  first_touch_date = min(date),
                  last_touch_date = max(date),
                  tot_time_lapse = round(as.numeric(last_touch_date - first_touch_date)),
                  conversion = sum(conversion)) %>%
        ungroup()
 
# distribution plot
ggplot(df_multi_paths_tl %>% filter(conversion == 1), aes(x = tot_time_lapse)) +
        theme_minimal() +
        geom_histogram(fill = '#4e79a7', binwidth = 1)
 
# cumulative distribution plot
ggplot(df_multi_paths_tl %>% filter(conversion == 1), aes(x = tot_time_lapse)) +
        theme_minimal() +
        stat_ecdf(geom = 'step', color = '#4e79a7', size = 2, alpha = 0.7) +
        geom_hline(yintercept = 0.95, color = '#e15759', size = 1.5) +
        geom_vline(xintercept = 23, color = '#e15759', size = 1.5, linetype = 2)
```



assuming reporting date as of January 10, 2016
```{r}
### for generic probabilistic model ###
df_multi_paths_tl_1 <- melt(df_multi_paths_tl[c(1:50), ] %>% select(customer_id, first_touch_date, last_touch_date, conversion),
                    id.vars = c('customer_id', 'conversion'),
                    value.name = 'touch_date') %>%
        arrange(customer_id)
rep_date <- as.Date('2016-01-10', format = '%Y-%m-%d')
 
ggplot(df_multi_paths_tl_1, aes(x = as.factor(customer_id), y = touch_date, color = factor(conversion), group = customer_id)) +
        theme_minimal() +
        coord_flip() +
        geom_point(size = 2) +
        geom_line(size = 0.5, color = 'darkgrey') +
        geom_hline(yintercept = as.numeric(rep_date), color = '#e15759', size = 2) +
        geom_rect(xmin = -Inf, xmax = Inf, ymin = as.numeric(rep_date), ymax = Inf, alpha = 0.01, color = 'white', fill = 'white') +
        theme(legend.position = 'bottom',
              panel.border = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              axis.ticks.x = element_blank(),
              axis.ticks.y = element_blank()) +
        guides(colour = guide_legend(override.aes = list(size = 5)))
```

You can see that id10072 finished the path in a conversion so we can add its retrospective touchpoints into the model. On the other hand, id10010‘s, id1001‘s and id10005‘s paths are fruitless as of reporting date but customer id10010 will purchase on January 19, 2016, customer id1001 will contact with a marketing channel on January 15, 2016, but won’t purchase and customer id10005 won’t have any new contacts with marketing channels, it is fruitless.

Therefore, our task is to compute generic model trying to identify which paths are completed as of reporting date both in a conversion or not. For example, we should use paths of id10072 and id10005 customers for computing the model, because we don’t expect new contacts or first purchases from them anymore.

We need to develop some criteria for identifying if an exact path has a higher chance to finish in a conversion or not. Again, we can analyze stats that characterize successful paths and make the assumption that if a path violates common stats towards success then it is fruitless with high probability.

There are at least two values I can suggest to start with:

1) time lapse from the first contact,

2) time lapse between the conversion date and a previous contact.

These values can be used as a combination of rules. From the above analysis we know that 95% of purchases were done in 23 days and we can compute time lapses between the conversion date and the previous contact

```{r}
df_multi_paths_tl_2 <- df_path_1_clean_multi %>%
        group_by(customer_id) %>%
        mutate(prev_touch_date = lag(date)) %>%
        ungroup() %>%
        filter(conversion == 1) %>%
        mutate(prev_time_lapse = round(as.numeric(date - prev_touch_date)))
         
# distribution
ggplot(df_multi_paths_tl_2, aes(x = prev_time_lapse)) +
        theme_minimal() +
        geom_histogram(fill = '#4e79a7', binwidth = 1)
 
# cumulative distribution
ggplot(df_multi_paths_tl_2, aes(x = prev_time_lapse)) +
        theme_minimal() +
        stat_ecdf(geom = 'step', color = '#4e79a7', size = 2, alpha = 0.7) +
        geom_hline(yintercept = 0.95, color = '#e15759', size = 1.5) +
        geom_vline(xintercept = 12, color = '#e15759', size = 1.5, linetype = 2)
```

We can see that 95% of customers have made a purchase within 12 days from the previous contact. Therefore, we can assume that if a customer made contact with a marketing channel the first time for more than 23 days and/or hasn’t made contact with a marketing channel for the last 12 days, then it is a fruitless path.

```{r}
# extracting data for generic model
df_multi_paths_tl_3 <- df_path_1_clean_multi %>%
        group_by(customer_id) %>%
        mutate(prev_time_lapse = round(as.numeric(date - lag(date)))) %>%
        summarise(path = paste(channel, collapse = ' > '),
                  tot_time_lapse = round(as.numeric(max(date) - min(date))),
                  prev_touch_tl = prev_time_lapse[which(max(date) == date)],
                  conversion = sum(conversion)) %>%
        ungroup() %>%
        mutate(is_fruitless = ifelse(conversion == 0 & tot_time_lapse > 20 & prev_touch_tl > 10, TRUE, FALSE)) %>%
        filter(conversion == 1 | is_fruitless == TRUE)
```






Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
